Setting up Data

### 2020-02-25 src/SetupData.Rmd
Finally debugged the kniting part, used here()

-rm Peromycus data BivData!!

@@

```{r setup, include=FALSE}
library(plyr)
library(dplyr)
library(raster)#for cV
library(ggplot2)
library(knitr)
library(kableExtra)
setwd("~./MLH1repo/")
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

#forgot why I'm reading in this metadata file
meta.data = read.csv("data/clean.Meta.Data.txt",sep = "\t", header=TRUE)
#remove all MAD (peromyscus) from meta.data
meta.data <- meta.data[ !grepl("MAD", meta.data$mouse) , ]

MLH1_data = read.csv("data/MLH1/AnonData.csv", header=TRUE)

knitr::opts_chunk$set(echo = FALSE)
```

@@

original_DF <- add_mouse(original_DF)
original_DF <- add_strain(original_DF)
original_DF <- add_category(original_DF)
original_DF <- add_sex(original_DF)

@@

@@

@@

```{r mouse.table, echo=FALSE, warning=FALSE}

#remove juvi mice

#Make a mouse level table -- why are there many old mice without any MLH1 (bc these are mice which were dissected, but not imaged/quantified)

AP_mouse_table.MLH1_HQ <- ddply(MLH1_data[MLH1_data$quality < 5,], c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(adj_nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(adj_nMLH1.foci),
                        var = as.numeric(format(round(   var(adj_nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(adj_nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
                        #quality?
)

@@

#add DOB to this table?
mice.frm.meta <- AP_mouse_table.MLH1_HQ[(AP_mouse_table.MLH1_HQ$mouse %in% meta.data$mouse),]

@@

#during merging -- try to just merge a few cols: 
AP_mouse_table_w.Ages <- merge(AP_mouse_table.MLH1_HQ, meta.data, by.y = "mouse", by.x = "mouse", all = TRUE)
#what are the differences between these two

@@

#remove cols with bad suffix
AP_mouse_table_w.Ages <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x, strain.y,
                    mouse.1  )  )

@@

#make sure they are all factors
AP_mouse_table.HQ$mouse <- as.factor(AP_mouse_table.HQ$mouse)
AP_mouse_table.HQ$sex <- as.factor(AP_mouse_table.HQ$sex)
AP_mouse_table.HQ$strain <- as.factor(AP_mouse_table.HQ$strain)
AP_mouse_table.HQ$subsp <-  as.factor(AP_mouse_table.HQ$subsp)

@@

@@

```{r HetC.DF.4MM, echo=FALSE}

@@

#use the merged DF to filter by age 
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
#main file, make everything unordered factors
AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)

DF.HetC.MixedModel <- AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$species == "M.musculus",]

#remove the extra merged cols
DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
                            subsp.y,subsp.x,
                    category.y,category.x, strain.x,
                    strain.y, mouse.1  )  )

#REMOVE STRAINS WITHOUT FEMALES
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
#remove musc strains
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]

DF.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used

#Only us the righ age window
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  (DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles

@@

#main df with variables coded as chr
DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)

#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))

#only female male reps

@@

#AP_mouse_table_w.Ages
#use the above table to pull the male mouse levels
male_mouse.av <-  AP_mouse_table_w.Ages %>% filter(sex== "male")

@@

male_mouse_table_w.Ages <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex.x == "male") &(AP_mouse_table_w.Ages$category.x != "other"),]

all.ages_plot <- ggplot(AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex.x == "male",], aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category.x)+ggtitle("Male Age Distributions")
all.ages_plot

@@

save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_2.2.20.RData")

>>>>

- rm Peromycus data BivData!!

@@

- try coding i) date image for each cell image, ii) most recent image date for each mouse

```{r setup, include=FALSE}
library(plyr)
library(dplyr)
library(raster)#for cV
library(ggplot2)
library(knitr)
library(kableExtra)
library(here)

#knitr::opts_chunk$set(echo = FALSE)

#opts_knit$set(root.dir = "~./MLH1repo/")
#setwd("~./MLH1repo/")

meta.data = read.csv(here("data", "/clean.Meta.Data.txt"),sep = "\t", header=TRUE)

#remove all MAD (peromyscus) from meta.data
meta.data <- meta.data[ !grepl("MAD", meta.data$mouse) , ]

MLH1_data = read.csv(here("data", "/MLH1/AnonData.csv"), header=TRUE)

source(here("src", "/CommonFunc_MLH1repo.R") )

@@

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
original_DF <- add_mouse(original_DF)
original_DF <- add_strain(original_DF)
original_DF <- add_category(original_DF)
original_DF <- add_sex(original_DF)

@@

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

@@

```{r Date.imaged, echo=FALSE}

#for issues with the total SC metrics, code in date imaged
#use the add_mouse function

#mini <- MLH1_data[1:500,]

count=1
for(i in MLH1_data$Original.Name){
   # print(i)
    templist= strsplit(i, split="_")[[1]]
    c = (templist[1])
    MLH1_data$image.string[count] <-  c
    #mini$image.date[count] <-  as.Date(c, format='%d%b%y')  #"%d%b%y")
    count= count +1
}
#works better if dat is converted out of image
MLH1_data$date.imaged <- as.Date(MLH1_data$image.string,  format='%d%b%y')

#MLH1_data <- add_mouse2(MLH1_data)

mini.MLH1_datamouse.table <- ddply(MLH1_data, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Random.Name),
                        #most recent date
                        oldest = min(date.imaged),
                        newest = max(date.imaged)

)

@@

```{r mouse.table, echo=FALSE}
#Make a mouse level table

AP_mouse_table.MLH1_HQ <- ddply(MLH1_data[MLH1_data$quality < 5,], c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(adj_nMLH1.foci),
                        mean_co = as.numeric(format(round(  mean(adj_nMLH1.foci, na.rm=TRUE), 3 ), nsmall=3) ),
                        cV = cv(adj_nMLH1.foci, na.rm=TRUE),
                        var = as.numeric(format(round(   var(adj_nMLH1.foci, na.rm=TRUE),3), nsmall=3)),
                        sd   = round(sd(adj_nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                        oldest.image.date = min(date.imaged),
                        newest.image.date = max(date.imaged)
                        #quality?
)

#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

@@

#add DOB to this table?
mice.frm.meta <- AP_mouse_table.MLH1_HQ[as.character(AP_mouse_table.MLH1_HQ$mouse) %in% as.character(meta.data$mouse),]

@@

#during merging -- try to just merge a few cols: 
AP_mouse_table_w.Ages <- merge(AP_mouse_table.MLH1_HQ, meta.data)
#don't use by, that makes duplicate cols
# by.y = "mouse", by.x = "mouse", all = TRUE)
#what are the differences between these two

@@

#remove cols with bad suffix
#AP_mouse_table_w.Ages <- subset(AP_mouse_table_w.Ages, select= -c(sex.y,sex.x,
#                            subsp.y,subsp.x,
#                    category.y,category.x, strain.x, strain.y,
#                    mouse.1  )  )

@@

#make sure they are all factors - this was breaking the kniting
#AP_mouse_table.HQ$mouse <- as.factor(AP_mouse_table.HQ$mouse)
#AP_mouse_table.HQ$sex <- as.factor(AP_mouse_table.HQ$sex)
#AP_mouse_table.HQ$strain <- as.factor(AP_mouse_table.HQ$strain)
#AP_mouse_table.HQ$subsp <-  as.factor(AP_mouse_table.HQ$subsp)

@@

#remove juvi mice via the list

@@

```{r HetC.DF.4MM, echo=FALSE}
#this block has so many fing errors

@@

#use the merged DF to filter by age 
#AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
#AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
#AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x
#main file, make everything unordered factors

#setwd("~./MLH1repo/src/")

source(here("src","CommonFunc_MLH1repo.R") )
AP_mouse_table_w.Ages <- add_species(AP_mouse_table_w.Ages)
DF.HetC.MixedModel <- AP_mouse_table_w.Ages %>% filter(species == "M.musculus")

#remove the extra merged cols
#DF.HetC.MixedModel <- subset(DF.HetC.MixedModel, select= -c(sex.y,sex.x,
#                            subsp.y,subsp.x,
#                    category.y,category.x, strain.x,
#                    strain.y, mouse.1  )  )

#REMOVE STRAINS WITHOUT FEMALES
#DF.HetC.MixedModel <- DF.HetC.MixedModel %>% filter( subsp != "Cast") %>% filter(strain != "AST")

# These are removed in the analysis
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$subsp != "Cast",]
#remove musc strains
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "AST",]
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "TOM",]
#remove PERC, MOLF, CZECH, (no females)
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "CZECH",]
#DF.HetC.MixedModel <- DF.HetC.MixedModel[DF.HetC.MixedModel$strain != "PERC",]
#F.HetC.MixedModel <- droplevels(DF.HetC.MixedModel)#remove levels not used

#Only us the righ age window
#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[( (DF.HetC.MixedModel$age.weeks < 25) &  #(DF.HetC.MixedModel$age.days != 13) ) | is.na(DF.HetC.MixedModel$age.weeks), ] #keeps mice without ages, and removes juvinilles

@@

#main df with variables coded as chr
#DF.HetC.MixedModel.HQ$subsp <- factor(DF.HetC.MixedModel.HQ$subsp, ordered = FALSE)
#DF.HetC.MixedModel.HQ$sex <- factor(DF.HetC.MixedModel.HQ$sex, ordered = FALSE)
#DF.HetC.MixedModel.HQ$strain <- factor(DF.HetC.MixedModel.HQ$strain, ordered = FALSE)
#DF.HetC.MixedModel.HQ$species <- factor(DF.HetC.MixedModel.HQ$species, ordered = FALSE)

#kable(table(DF.HetC.MixedModel.HQ$sex, DF.HetC.MixedModel.HQ$strain))
#only female male reps

@@

#AP_mouse_table_w.Ages
#use the above table to pull the male mouse levels
mlh1male_mouse.av <-  AP_mouse_table_w.Ages %>% filter(sex== "male")

@@

male_mouse_table_w.Ages <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$sex == "male") &(AP_mouse_table_w.Ages$category != "other"),]

all.ages_plot <- ggplot(AP_mouse_table_w.Ages[AP_mouse_table_w.Ages$sex == "male",], aes(age.weeks))+geom_histogram(binwidth = 3)+facet_wrap(~category)+ggtitle("Male Age Distributions")
all.ages_plot

@@

#remove non relevant files -- not sure where they load from
#BD data and Peromyscus data
#rm(c(BD_MLH1_data, BDmouse_level_table, MLH1_data.hq) )

save.image("~./MLH1repo/data/MLH1/MLH1_data_setup_2.24.20.RData")
